### 2022-07-05 iclfs_review/lit_review.Rmd
Add literature review markdown file

---
title: The aspects of integrated crop-livestock-forest system in Brazil

author: Hugo Tameirão Seixas

output:
  html_document:
    theme: readable
    number_sections: true

---

# Title and Abstract

# Aims and questions

# Review registration

# Eligibility criteria

# Finding studies

## Define the type of search (e.g.comprehensive search, representative sample)

## State what sources of information were sought (e.g. published and unpublished studies, personal communications)

## Include, for each database searched, the exact search strings used, with keyword combinations and Boolean operators

Scopus: TITLE-ABS-KEY(iclf OR iclfs OR crop-livestock-forestry OR crop-livestock-forest OR agrosilvipastoral OR ilpf OR lavoura-pecuária-floresta OR agrosilvipastoril)

Embrapa: (ICLF) OR (iclfs) OR (crop-livestock-forestry) OR (crop-livestock-forest) OR (agrosilvipastoral) OR (ilpf) OR (lavoura-pecuária-floresta) OR (agrosilvipastoril)

## Provide enough information to repeat the equivalent search (if possible), including the timespan covered (start and end dates)

# Study selection

# Data collection process

# Data items

# Assessment of individual study quality

# Effect size measures

# Missing data

# Meta-analytic model description

# Software

# Non independence

# Meta-regression and model selection

# Publication bias and sesitivity analysis

# Clarification of post hoc analysis

# Metadata data and code

# Results of study selection process

# Sample sizes and study characteristics

# Meta-analysis

# Heterogeneity

# Meta-regression

# Outcomes of publication bias and sensitivity analyses

# Discussion

# Contributions and funding

# References