### 2022-07-07 iclfs_review/lit_review.Rmd
Wrote some aims of the review

---
title: The aspects of integrated crop-livestock-forest system in Brazil

author: Hugo Tameirão Seixas

output:
  html_document:
    theme: readable
    number_sections: true

---

# Title and Abstract

# Aims and questions

## Provide a rationale for the review

## Reference any previous reviews or meta-analyses on the topic

## State the aims and scope of the review (including its generality)

In this review I will analyse studies about ICLFS in Brazil. I pretend to answer several questions:

- Where are experiments on ICLFS being conducted?

- What are the characteristics of these systems? (species used, fertilizers, number of individuals per hectare, duration of the production cycle...)

- What are the fiber and food productivity in these systems?

- What is the amount of above ground carbon (alive and dead) being accumulated?

- What is the amount of below ground carbon being accumulated?

- What are the emissions of green house gases related to the ICLFS?

- What is the economic revenue from the adoption of the ICLFS?

## State the primary questions the review addresses (e.g. which moderators were tested)

## Describe whether effect sizes were derived from experimental and/or observational comparisons

# Review registration

# Eligibility criteria

# Finding studies

## Define the type of search (e.g.comprehensive search, representative sample)

## State what sources of information were sought (e.g. published and unpublished studies, personal communications)

## Include, for each database searched, the exact search strings used, with keyword combinations and Boolean operators

Scopus: TITLE-ABS-KEY(iclf OR iclfs OR crop-livestock-forestry OR crop-livestock-forest OR agrosilvipastoral OR ilpf OR lavoura-pecuária-floresta OR agrosilvipastoril)

Embrapa: (ICLF) OR (iclfs) OR (crop-livestock-forestry) OR (crop-livestock-forest) OR (agrosilvipastoral) OR (ilpf) OR (lavoura-pecuária-floresta) OR (agrosilvipastoril)

## Provide enough information to repeat the equivalent search (if possible), including the timespan covered (start and end dates)

# Study selection

# Data collection process

# Data items

# Assessment of individual study quality

# Effect size measures

# Missing data

# Meta-analytic model description

# Software

# Non independence

# Meta-regression and model selection

# Publication bias and sesitivity analysis

# Clarification of post hoc analysis

# Metadata data and code

# Results of study selection process

# Sample sizes and study characteristics

# Meta-analysis

# Heterogeneity

# Meta-regression

# Outcomes of publication bias and sensitivity analyses

# Discussion

# Contributions and funding

# References