### 2022-07-25 iclfs_review/lit_review.qmd
Create literature review document using Quarto.

---
title: "Characterization of integrated crop-livestock-forest system in Brazil"
author: "Hugo Tameirão Seixas"
format: 
  html:
    page-layout: article
editor: visual
---

## Title and Abstract

## Aims and questions

### Provide a rationale for the review

### Reference any previous reviews or meta-analyses on the topic

### State the aims and scope of the review (including its generality)

In this review I will analyse studies about ICLFS in Brazil. I pretend to answer several questions:

-   Where are experiments on ICLFS being conducted?

-   What are the characteristics of these systems? (species used, fertilizers, number of individuals per hectare, duration of the production cycle...)

-   What are the fiber and food productivity in these systems?

-   What is the amount of above ground carbon (alive and dead) being accumulated?

-   What is the amount of below ground carbon being accumulated?

-   What are the emissions of green house gases related to the ICLFS?

-   What is the economic revenue from the adoption of the ICLFS?

### State the primary questions the review addresses (e.g. which moderators were tested)

### Describe whether effect sizes were derived from experimental and/or observational comparisons

## Review registration

## Eligibility criteria

## Finding studies

### Define the type of search (e.g.comprehensive search, representative sample)

### State what sources of information were sought (e.g. published and unpublished studies, personal communications)

### Include, for each database searched, the exact search strings used, with keyword combinations and Boolean operators

Scopus: TITLE-ABS-KEY(iclf OR iclfs OR crop-livestock-forestry OR crop-livestock-forest OR agrosilvipastoral OR ilpf OR lavoura-pecuária-floresta OR agrosilvipastoril)

Embrapa: (ICLF) OR (iclfs) OR (crop-livestock-forestry) OR (crop-livestock-forest) OR (agrosilvipastoral) OR (ilpf) OR (lavoura-pecuária-floresta) OR (agrosilvipastoril)

### Provide enough information to repeat the equivalent search (if possible), including the timespan covered (start and end dates)

## Study selection

## Data collection process

## Data items

## Assessment of individual study quality

## Effect size measures

## Missing data

## Meta-analytic model description

## Software

## Non independence

## Meta-regression and model selection

## Publication bias and sesitivity analysis

## Clarification of post hoc analysis

## Metadata data and code

## Results of study selection process

## Sample sizes and study characteristics

## Meta-analysis

## Heterogeneity

## Meta-regression

## Outcomes of publication bias and sensitivity analyses

## Discussion

## Contributions and funding

## References