### 2022-09-23 manuscript/manuscript.qmd
Add text to the methods section.

---
title: Integrated farming and global warming mitigation in Brazil
author: 
  - name: Hugo Tameirão Seixas
    email: seixas.hugo@protonmail.com
    url: hugotseixas.com
    affiliations:
      - name: State University of Campinas
toc: true
toc-title: Contents
format: 
  html:
    theme: lux
page-layout: article
bibliography: references.bib
link-citations: true
abstract: "There is a high expectation in the reduction of liquid greenhouse gases in Brazil by the implementation of integrated crop-livestock-forestry systems. The foundation of such expectation relies in the hypothesis that these systems are capable of storing more above and below ground carbon, in relation to conventional farming systems, due to the presence of trees. However, that are not robust estimates about the potential of these systems to store carbon in a national scale, and usually point estimates are used to extrapolate carbon sequestration. In order to improve the representation of carbon sequestration by integrated crop-livestock-forestry systems, this project aims to use published data to develop a spatially explicit model to estimate carbon fluxes in Brazil."
appendix-style: plain
license: CC BY-NC-ND 4.0
---

```{r}
#| include: false

library(DiagrammeR)
library(lubridate)
library(tidyverse)

```

------------------------------------------------------------------------

# Introduction

The implementation of integrated crop-livestock-forestry (ICLF) systems in Brazil are expected to sequester enough carbon to contribute to reduce annual liquid carbon emissions. Despite the expectations, there are not clear estimates of the magnitude of reductions of liquid emissions caused by ICLF expansion. Field studies show contrasting results. @Valani2022 observed a higher organic carbon content in the soil in relation to conventional grazing and other integrated systems. @Bieluczyk2020 field measurements did not show clear differences between extensive grazing and ICLF soil organic matter, with highly temporal heterogeneity in observations. @Silva2018837 found inconsistent differences of soil total organic carbon between pasture and ILPF system due to high heterogeneity of observations along the soil layers, and observed small differences between agriculture and integrated system, in which the agriculture presented higher carbon stocks. @Oliveira2018105 found higher amounts of soil organic carbon of two ICLF systems and continuous pasture areas, there were also observed differences between the two integrated systems.

Unfortunately, estimates of carbon sequestration by ICLF at larger scales usually use point estimates, which can severely bias the results towards one direction. @Azevedo_2018 used a CO<sub>2</sub>-eq removal factor of -6.24 t CO<sub>2</sub>-eq ha<sup>-1</sup> y<sup>-1</sup>, which was extracted from one experiment comparing integrated crop-livestock (ICL) systems to pasture [@Carvalho_2010], and did not observe consistent differences between these types of systems. The removal factor was then extrapolated over estimated ILPF area, this simplification is the best guess that Brazil have over the potential of integrated systems in reducing liquid emissions. In face of such scenario, this study aims to generate estimations of CO<sub>2</sub>-eq removal accounting for variation in many aspects, soil attributes, ICLF system characteristics, species used, spacing and climate. We expect that such estimations will provide more realistic ranges of possible outcomes in a diverse set of future scenarios of ICLF expansion in Brazil.

------------------------------------------------------------------------

# Objectives

The present work is composed by three main objectives, accompanied by parallel related questions:

1. Identify ICLF systems characteristics and dynamics in Brazil;

    1.1. What are the management characteristics of ICLF systems?

    1.2. What are the compositions of each component of ICLF systems?

    1.3. What is the productivity of each component of ICLF systems?

    1.4. What are the effects of ICLF system over soil properties?

    1.5. What is the economical revenue of ICLF systems?

    1.6. What is the potential expansion of ICLF systems in Brazil?

2. Estimate the potential of ICLF systems to provide benefits, in the context of climate change;

    2.1. What are the ICLF system management characteristics that provide the best balance between carbon removal, productivity and financial revenue?

    2.2. What are the main factors that influence the carbon removal in the modelling framework?

    2.3. What is the pottential of the expansion of ICLF systems over the emissions reduction pledges made by Brazil?

3.  Evaluate the implementation of ICLF as a mean to curb and mitigate climate change, in relation to different perspectives of solutions to avoid climatic catastrophe;

    3.1. Does the expansion of ICLF systems taking into account equity and possible social impacts, as a public policy?

    3.2. The actual strategies to promote the adoption of ICLF systems are being effective?

    3.3. Does focusing on lowering net emissions a distraction from lowering gross emissions?

------------------------------------------------------------------------

# Materials and Methods

In order to achieve the proposed objective, the work will be divided in three major parts:

1. Overview of ICLF characteristics in Brazil;

2. Future projections and estimates of benefits generated by ICLF systems adoption;

3. Critical analysis of ICLF as a public policy to curb climate change;

```{r}
#| echo: false

grViz(
  "digraph {

  graph [layout = dot, 
         rankdir = TD, 
         nodesep = 2, 
         ranksep = 0.5,
         tooltip = ' ']

  node [shape = box,
        fontname = Helvetica,
        style = filled,
        fontcolor = '#343a40'
        fillcolor = '#f7f7f9'
        penwidth = 2.5,
        width = 3,
        height = 1.2,
        fontsize = 25,
        tooltip = ' ']

  edge [penwidth = 2.5, 
        arrowsize = 2,
        tooltip = ' ']

  a [label = 'Systematic \n Review', xlabel = '1']
  b [label = 'Experts Survey', xlabel = '1']
  c [label = 'ICLF overview \n data set', xlabel = '1']
  d [label = 'Projections \n Scenarios', xlabel = '2']
  e [label = 'LULC \n Projections', xlabel = '2']
  f [label = 'Benefits \n Estimations', xlabel = '2']
  g [label = 'Targeted \n Review', xlabel = '3']
  h [label = 'Narrative \n Discussion', xlabel = '3']

  a -> c b -> c
  c -> d d -> e
  e -> f c -> f
  f -> h g -> h

  }"
)

```

### Integrated crop-livestock-forest overview

The overview of ICLF systems in Brazil will be performed by analyzing published research and web surveys with experts.

#### Literature review

The literature review will be performed over published research from two sources, Scopus and Embrapa. It will have a very general approach, in which any research analyzing any aspect of ICLF will be screened. The strings that will be used to search research are the following:

*   Scopus: TITLE-ABS-KEY(iclf OR iclfs OR crop-livestock-forestry OR crop-livestock-forest OR agrosilvipastoral OR ilpf OR lavoura-pecuária-floresta OR agrosilvipastoril) AND PUBYEAR > 1999

* Embrapa: ((ICLF) OR (iclfs) OR (crop-livestock-forestry) OR (crop-livestock-forest) OR (agrosilvipastoral) OR (ilpf) OR (lavoura-pecuária-floresta) OR (agrosilvipastoril)) AND (ano-publicacao:[2000 TO *]) AND (tipo-material-sigla:"PC" OR tipo-material-sigla:"AP" OR tipo-material-sigla:"NC" OR tipo-material-sigla:"RT") AND (idioma:"Inglês" OR idioma:"Português")

The data and information extract from selected research will be diverse, and will be organized in 11 [relational tables](../iclfs_review/tables/00_tables_collection.html). The tables will cover general information about the research document and characteristics of the experiment, as also data about the system management, soil properties, productivity, growth, emissions, and radiation. All tables will be described by metadata files.

An important assumption of the review is that the ICLF experiments are representative of the ICLF being implemented by farmers in Brazil.

#### Experts survey

The experts survey will be performed by a web application where the participants will be able to interact with a questionnaire to provide information about ICLF systems.

The nature of the survey is semi-structured, with closed and open questions. The participants will have the option to avoid answering any question.

The participants desired profile is of researchers or technicians with any experience with ICLF systems. A minimum of 50 participants is expected.

The objective of the survey is to gather information that are not available in the literature, as also to provide expected scenarios to the ICLF expansion in the future.

All answers will be stored in tables for posterior analysis. Each question will be represented by a column, and each participant will be represented as a row (one observation for each question).

The web application will be deployed in a personal website, where the participants will receive the link of the survey, along with a token that will be used to submit the answers.

### Model projections

Model projections will estimate emissions and removals of Greenhouse Gases (GHG), food and fiber productivity and income of ICLF systems. The projections will be performed by a mixed modelling framework, using random processes, empirical and mechanistic models, deterministic and probabilistic factors to estimate the variables of interest.

The projections will cover the whole Brazilian territory, from 2020 to 2100. Estimations will be made over a grid formed by 3x3 kilometers cells.

The model will assume that ICLF will only occupy areas considered as degraded pasture according to the classification performed by MapBiomas [@Souza_2020]. Another assumption is that after a transition from degraded pasture to ICLF, it will never change to another use.

#### Projections scenarios

To conduct the projections, a set of possible scenarios will be created. The scenarios will be based on data and information gathered from the systematic review and the experts survey.

The ICLF expansion scenarios will be based on the expectations of experts, and will provide data to create three scenarios. One will be the median of the expected area to be occupied by ICLF in 2030, 2050 and 2100, the other two will be the upper and lower quartiles. These three scenarios will be considered as reasonable future projections. There are going to be other two extreme scenarios, one where there wont be any expansion of ICLF, and another where the ICLF will cover all the degraded pasture extent by 2100.

The ICLF management scenarios will be based on the systematic review and experts survey. The management scenarios will dictate the characteristics of the ICLF systems being implemented in the future, as the species, population, spacing, inputs, and cycle length. The management scenarios will depend on the region where the ICLF system is being implemented.

```{r}
#| echo: false

grViz(
  "digraph {

  graph [layout = dot,
         rankdir = TD, 
         nodesep = 2, 
         ranksep = 0.5,
         tooltip = ' ']

  node [shape = box,
        fontname = Helvetica,
        style = filled,
        fontcolor = '#343a40'
        fillcolor = '#f7f7f9'
        penwidth = 2.5,
        width = 3,
        height = 1.2,
        fontsize = 25,
        tooltip = ' ']

  edge [penwidth = 2.5, 
        arrowsize = 2,
        tooltip = ' ']

  a [label = 'Region']
  b [label = 'Species']
  c [label = 'Population']
  d [label = 'Cycle Length']
  e [label = 'Management \n Practices']

  a -> b b -> c
  b -> d b -> e

  }"
)

```

#### Land use projections

The land use projections will be performed by randomly assigning grid cells occupied by degraded pastures to be converted to ICLF. The assigned grid cell will first be occupied by a transition state, which will be a Integrated Crop Livestock (ICL) system, and after a certain period, it will finally change to a ICLF. The assigned cells will also be attributed with management characteristics from the management scenarios explained in the section above.

The rate of ICLF expansion from year to year will be closely linear, and will be dictated by the ICLF expansion scenario.

The ICLF expansion process will be repeated 100 times for each ICLF expansion scenario, providing a range of outcomes in the final results.

```{r}
#| echo: false

grViz(
  "digraph {

  graph [layout = dot,
         rankdir = LR, 
         nodesep = 1.2, 
         ranksep = 0.5,
         tooltip = ' ']

  node [shape = box,
        fontname = Helvetica,
        style = filled,
        fontcolor = '#343a40'
        fillcolor = '#f7f7f9'
        penwidth = 2.5,
        width = 3,
        height = 1.2,
        fontsize = 25,
        tooltip = ' ']

  edge [penwidth = 2.5, 
        arrowsize = 2,
        tooltip = ' ']

  a [label = 'Pasture']
  b [label = 'Transition \n (ICL)']
  c [label = 'ICLF']

    node [shape = oval,
        fontname = Helvetica,
        style = filled,
        fontcolor = '#343a40'
        fillcolor = '#f7f7f9'
        penwidth = 2.5,
        width = 3,
        height = 1.2,
        fontsize = 25,
        tooltip = ' ']

  e [label = 'Random Variables']
  f [label = 'Expansion \n Scenario']

      node [shape = diamond,
        fontname = Helvetica,
        style = filled,
        fontcolor = '#343a40'
        fillcolor = '#f7f7f9'
        penwidth = 2.5,
        width = 3,
        height = 1.2,
        fontsize = 25,
        tooltip = ' ']

  d [label = 'Random \n Selection']

  a -> d b -> c c -> b 
  d -> b e -> b e -> c
  f -> d

  }"
)

```

#### Benefits estimations

```{r}
#| echo: false

grViz(
  "digraph {

  graph [layout = dot,
         rankdir = TD, 
         nodesep = 1.2, 
         ranksep = 0.5,
         tooltip = ' ']

  node [shape = box,
        fontname = Helvetica,
        style = filled,
        fontcolor = '#343a40'
        fillcolor = '#f7f7f9'
        penwidth = 2.5,
        width = 3,
        height = 1.2,
        fontsize = 25,
        tooltip = ' ']

  edge [penwidth = 2.5, 
        arrowsize = 2,
        tooltip = ' ']

  a [label = 'Biomass']
  b [label = 'Soil Carbon \n (Century)']
  c [label = 'Productivity']
  d [label = 'Revenue']

    node [shape = oval,
        fontname = Helvetica,
        style = filled,
        fontcolor = '#343a40'
        fillcolor = '#f7f7f9'
        penwidth = 2.5,
        width = 3,
        height = 1.2,
        fontsize = 25,
        tooltip = ' ']

  e [label = 'Population']      
  f [label = 'Climate']
  g [label = 'Species']
  h [label = 'Management \n Practices']

  a -> b a -> c c -> d

  e -> a e -> c
  f -> a f -> b f -> c
  g -> a g ->b g -> c
  h -> d

  }"
)

```

### Critical analysis

------------------------------------------------------------------------

# Results

------------------------------------------------------------------------

# Discussion

------------------------------------------------------------------------

# Conclusions

------------------------------------------------------------------------

# Schedule

```{r}
#| echo: false

schedule_table <- 
  bind_rows(
    list(
      activity = "Systematic Review",
      start = "01-10-2022",
      end = "01-06-2023",
      section = 1
    ),
    list(
      activity = "Experts Survey Development",
      start = "01-10-2022",
      end = "01-04-2023",
      section = 1
    ),
    list(
      activity = "Experts Survey Application",
      start = "01-04-2023",
      end = "01-08-2023",
      section = 1
    ),
    list(
      activity = "Model Prototype Development",
      start = "01-10-2022",
      end = "01-09-2023",
      section = 2
    ),
    list(
      activity = "Projection Scenarios Development",
      start = "01-07-2023",
      end = "01-09-2023",
      section = 2
    ),
    list(
      activity = "LULC Projections",
      start = "01-09-2023",
      end = "01-11-2023",
      section = 2
    ),
    list(
      activity = "Benefits Estimations",
      start = "01-11-2023",
      end = "01-04-2024",
      section = 2
    ),
    list(
      activity = "Targeted Review",
      start = "01-01-2024",
      end = "01-06-2024",
      section = 3
    ),
    list(
      activity = "Narrative Discussion",
      start = "01-04-2024",
      end = "01-08-2024",
      section = 3
    ),
    list(
      activity = "Final Manuscript Development",
      start = "01-08-2024",
      end = "01-02-2025",
      section = NA
    )
  )

schedule_table <- schedule_table %>%
  mutate(
    start = dmy(start),end = dmy(end),
    activity = factor(activity, levels = activity)
  ) %>%
  pivot_longer(
    cols = c(start, end),
    names_to = "when",
    values_to = "date"
  )

schedule_table %>%
    ggplot() +
    geom_line(
      aes(x = date, y = activity), 
      size = 3, 
      lineend = "round"
    ) +
    scale_y_discrete(limits = rev) +
    scale_x_date(
      date_labels = "%b-%Y", 
      date_breaks = "6 months", 
      date_minor_breaks = "1 month",
      limits = c(ymd("2022-09-01"), ymd("2025-03-01"))
    ) +
    labs(color = "Section") +
    theme_minimal() + 
    theme(
      text = element_text(family = "sans", size = 15),
      panel.grid.major = element_line(color = "#a9a9a9"),
      axis.title = element_blank(),
      axis.text.x = element_text(angle = 35, vjust = 0.6),
      plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
      legend.position = "bottom"
    )

```

------------------------------------------------------------------------

# Supporting Information

# References