### 2023-02-07 manuscript/lct_se_ac.qmd
Add content to the whole manuscript.

---
title: "Dimensions of the impacts of land transitions in the Amazon and Cerrado"
author: 
  - name: Hugo Tameirão Seixas
    email: seixas.hugo@protonmail.com
    url: hugotseixas.com
    affiliations:
      - name: State University of Campinas
format: 
  html:
    theme: lux
page-layout: article
bibliography: references.bib
link-citations: true
editor_options: 
  chunk_output_type: console
execute:
  freeze: auto
  cache: true
license: CC BY-NC-ND 4.0
---

```{r}
#| include: false

library(sf)
library(geobr)
library(gt)
library(DiagrammeR)
library(lubridate)
library(tidyverse)

```

## Introduction

## Objectives

The present work is composed of three main objectives, accompanied by related questions:

1.  Predict probability and area of native vegetation suppression in the Amazon and Cerrado;
    a.  What are the main drivers of native vegetation suppression?
    b.  How do drivers interact among them?
2.  Predict land cover transition trajectories after native vegetation suppression;
    a.  What are the main drivers of land cover dynamics after native vegetation suppression?
    b.  How do drivers interact among them?
    c.  What are the main patterns for future projections in the Amazon and Cerrado?
3.  Predict the impact of land cover transitions over environmental and social variables;
    a.  What is the importance of land cover transitions in predicting environmental and social variables?
    b.  Does land-use transitions interact with other drivers of environmental and social variables?
    c.  What are the expected social and environmental projections in different land-cover trajectories for the near future?

## Materials and Methods

To achieve the proposed goal, I will divide the work into three major parts:

1.  Literature overview of the main drivers of the expected outcomes (vegetation suppression, land cover transition, social and environmental variables);

2.  Screening for available data and development of structured data collection;

3.  Development of models;

4.  Future projections;

```{r}
#| echo: false

grViz(
  "digraph {

  graph [layout = dot, 
         rankdir = TD, 
         nodesep = 1.2, 
         ranksep = 0.8,
         tooltip = ' ']

  node [shape = box,
        fontname = Helvetica,
        style = filled,
        fontcolor = '#343a40'
        fillcolor = '#f7f7f9'
        penwidth = 2.5,
        width = 3,
        height = 1.2,
        fontsize = 25,
        tooltip = ' ']

  edge [penwidth = 2.5, 
        arrowsize = 2,
        tooltip = ' ']

  a [label = 'Literature \n Review', xlabel = '1']
  b [label = 'Experts Survey', xlabel = '1']
  c [label = 'Available Data', xlabel = '2']
  d [label = 'Database \n Structuring', xlabel = '2']
  e [label = 'Models \n Development', xlabel = '3']
  f [label = 'Projections \n Scenarios', xlabel = '4']
  g [label = 'Projections \n Estimations', xlabel = '4']

  a -> c a -> d
  c -> d b -> d
  b -> e b -> f
  e -> g f -> g
  d -> e

  }"
)

```

\

The literature review will provide a theoretical and practical basis for the modelling section.

The database structuring will store all the relevant data in a single file structure, so that the modelling section can be performed in a clean and organized manner.

The models will be developed following standard procedures and libraries that allow the implementation of different techniques. Each model will be stored as a mean to be reproduced when needed.

Future projections will serve as a bridge to discuss how public policies can be planned to mitigate negative impacts of native vegetation suppression in the Amazon and Cerrado.

#### Literature Review

The literature review will be performed over different subjects. For all of them, an hybrid approach will be adopted, in which database searches are combined with snowball searches.

```{r}
#| echo: false

grViz(
  "digraph {

  graph [layout = dot, 
         overlap = false,
         rankdir = TB, 
         nodesep = 5, 
         ranksep = 0.4,
         tooltip = ' ']

  node [shape = box,
        fontname = Helvetica,
        style = filled,
        fontcolor = '#343a40'
        fillcolor = '#f7f7f9'
        penwidth = 2.5,
        width = 3,
        height = 1.2,
        fontsize = 25,
        tooltip = ' ']

  edge [penwidth = 2.5, 
        arrowsize = 2,
        tooltip = ' ']

  a [label = 'Specific \n Questions']
  b [label = 'Query \n Development']
  c [label = 'Database Search']
  d [label = 'Documents \n Screening']
  e [label = 'Documents \n Collection']
  f [label = 'Forward Snowball \n Search']
  g [label = 'Final Collection']

  a -> b a -> d
  b -> c c -> d
  d -> e e -> f
  f -> d e -> g

  }"
)

```

\

The literature review will serve to answer questions about variables that will be predicted in this project. For each variable, I want to explore what are the main drivers that are causing them, how these drivers relate with each other, and what were the methods used to estimate the effects and interactions.

#### Database Structuring

The database will contain data from different themes.
To structure a collection of data which presents different nature, I will transform them into a common grid system.
The grid system will be composed by 0.3 degrees cells.

```{r}
#| echo: false
#| message: false

biomes <- read_biomes(showProgress = FALSE) %>%
  filter(name_biome %in% c("Amazônia", "Cerrado")) %>%
  st_make_valid()

grid <- 
  st_make_grid(biomes, cellsize = 0.3) %>%
  st_as_sf() %>%
  st_filter(biomes, .predicate = st_intersects)

ggplot() +
  geom_sf(data = biomes) +
  geom_sf(data = grid, fill = "transparent") +
  theme_void()

```

Each grid cell will contain an observation, for each year, for all the variables that will be included in the study.

#### Model Development

The models will follow a sequential order.
The output of a model will serve as a feature to the next model.

```{r}
#| echo: false

table <-
  tibble(
    section = c(
      "Vegetation suppression probability", "Vegetation suppression area", 
      "Land cover transitions", "Transitions impacts"
    ),
    model_type = c(
      "Classification", "Regression",
      "Classification", "Regression"
    )
  )

table %>%
  pivot_wider(names_from = section, values_from = model_type) %>%
  gt() %>%
  cols_align(align = "center") %>%
  tab_options(
    table.border.top.style = "hidden",
    table.border.bottom.style = "hidden"
  )

```

The first model will classify areas that suffered vegetation suppression.
So the outcome will be a binary result (0 = no suppression; 1 = suppression).
I want that this model to have the minimum inclusion errors for the class "1".
In other words, I want to avoid, as best as I can, classifying an area with label "1", when the reality is that the area should be label "0".
To achieve this, the model will be adjusted to present the minimum inclusion error for the class "1", and the probability threshold for class "1" will be a higher value.

The second model will predict the amount of vegetation suppression that was classified with the label "1" in the previous model.
The predictions will have to be bounded by the amount of natural vegetation inside the grid cell.

The third model will predict the predominant land use transitions of the areas that suffered native vegetation suppression.

The fourth model will be actually a set of models.
Each model will be responsible to predict environmental and social outcomes from land transition patterns.

## Results

The study will provide three main quantitative results, predictions of vegetation suppression, predictions of land use transitions, and predictions of social and environmental variables.

More than only the predictions, I will also analyse and present explanations of the models behavior, and inferences about the studied phenomena (will only take this leap if I feel confident with the models).

## Discussion

The discussion of the results will focus on two fronts, comparing our results with previous studies, and build a bridge with discussions from social sciences and political science.

## Conclusions

## Schedule

```{r}
#| echo: false
#| warning: false

schedule_table <- 
  bind_rows(
    list(
      activity = "Literature Review",
      start = "01-02-2023",
      end = "01-06-2023",
      section = 1
    ),
    list(
      activity = "Experts Survey",
      start = "01-06-2023",
      end = "01-01-2024",
      section = 1
    ),
    list(
      activity = "Data Screening",
      start = "01-02-2023",
      end = "01-08-2023",
      section = 2
    ),
    list(
      activity = "Database Structuring",
      start = "01-04-2023",
      end = "01-09-2023",
      section = 2
    ),
    list(
      activity = "Exploratory Data Analisys",
      start = "01-06-2023",
      end = "01-11-2023",
      section = 2
    ),
    list(
      activity = "Models Development",
      start = "01-11-2023",
      end = "01-05-2024",
      section = 3
    ),
    list(
      activity = "Projections Estimations",
      start = "01-05-2024",
      end = "01-09-2024",
      section = 4
    )
  )

schedule_table <- schedule_table %>%
  mutate(
    start = dmy(start),end = dmy(end),
    activity = factor(activity, levels = activity)
  ) %>%
  pivot_longer(
    cols = c(start, end),
    names_to = "when",
    values_to = "date"
  )

schedule_table %>%
    ggplot() +
    geom_line(
      aes(x = date, y = activity), 
      linewidth = 3, 
      lineend = "round"
    ) +
    scale_y_discrete(limits = rev) +
    scale_x_date(
      date_labels = "%b-%Y", 
      date_breaks = "6 months", 
      date_minor_breaks = "1 month",
      limits = c(ymd("2022-09-01"), ymd("2025-03-01"))
    ) +
    labs(color = "Section") +
    theme_minimal() + 
    theme(
      text = element_text(family = "sans", size = 15),
      panel.grid.major = element_line(color = "#a9a9a9"),
      axis.title = element_blank(),
      axis.text.x = element_text(angle = 35, vjust = 0.6),
      plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
      legend.position = "bottom"
    )

```

## References