### 2023-03-24 README.Rmd
Update documentation.

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# TICA: Dimensions of land transitions impacts in the Cerrado and Amazon

<!-- badges: start -->
<!-- badges: end -->

## Installation

You can install the development version of TICA from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("hugotseixas/tica")
```