### 2023-03-24 README.Rmd
Update readme.

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "/man/figures/README-",
  out.width = "100%"
)
```

# TICA: Dimensions of land transitions impacts in the Cerrado and Amazon <a href="https://hugotseixas.com/en/tica/"><img src="/man/figures/logo.png" align="right" height="138" /></a>

<!-- badges: start -->
<!-- badges: end -->

## Overview

TICA is a model being develop to estimate deforestation, land use transitions, and their impact on social and environmental aspects of the Cerrado and Amazônia biomes in Brazil. This model is part of an effort to develop an interdisciplinary approach to analyse these phenomena.

## Installation

You can install the development version of TICA from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("hugotseixas/tica")
```